Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-04-19 arcticgrids.tex
cleaned up citations

In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
{\color{blue}(WHL: Add refs)}
These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly. While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids 
{\color{blue}(define 'reduced grids')}
and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid. We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.

@@

The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which have large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}. Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c). Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure. {\color{blue} (WHL: Add 1 or 2 refs)} This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement. In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).  However, refinement comes at the cost of a smaller CFL-limited time step in the refined region. (The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.)

@@

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias (Pollard 2000) \cite{VETAL2018TC}. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low resolution GCMs, and may further degrade the simulated SMB. For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates (citations for MM5, RACMO and MAR). The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB. However, unlike GCMs, RCMs are not a freely evolving system and boundary conditions must be prescribed at the lateral boundaries. The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system.

@@

The SE dycore has been an available option since CESM, version 1, but has been under active development ever since. This includes the switch to a dry-mass vertical coordinate \cite{LetAl2018JAMES} and incorporation of an accelerated multi-tracer transport scheme \cite{LTOUNGK2017MWR}, made available in CESM2. This paper documents several additional enhancements to the SE dycore as part of the release of CESM2.2. These include three new VR configurations (Figure~\ref{fig:vr-grids}), two Arctic meshes and a Contiguous United-States mesh ($\tt CONUS$; featured in \citeA{GetAl2017JAMES,MUSCIA}). While there are many published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,bambach}, these studies either used development code or collaborated closely with model developers. CESM2.2 is the first code release that contains out of the box VR functionality in CESM.

This study compares the representation of high-latitude regions using the SE and FV dycores in CESM2.2 (see description below), as these two dycores treat high latitudes (i.e., the pole problem) in different ways. Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods. Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.

@@

The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{Gettelman2019GRL}, supports several different atmospheric dynamical cores. These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores {\color{green}{Andrew - maybe not mention EUL or FV3?}}.

[phl: maybe mention that FV was used for IPCC simulations; SE in iHESP]

@@

The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate. Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong]. A polar filter damps computational instability due to the convergence of meridians, allowing a longer time step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}. The form of the filter is designed to slow down the propagation of large zonal wave-numbers to satisfy the CFL condition of the shortest resolved wave at some reference latitude.

@@

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements to uniform-resolution grids: (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.

@@

We evaluate model simulations on six different grids in this study (Table~\ref{tbl:table1}). The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b). We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. We run an additional $1^{\circ}$ CAM-SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d; note CSLAM is still run on the $3\times3$ control volume grid;>{HETAL2019JAMES}.

Two VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region. The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.

@@

The physics time step depends on the grid resolution. Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $\tt Arctic$ and $\tt Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).

@@

All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{Gettelman2019GRL}. The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2,Morrison2015}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}. Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).

Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore. The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance. We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb_gamma$ is reduced from 0.308 to 0.270) and also reduced the strength of the damping for horizonatal component of turbulent energy ($\tt clubb_c14$ is reduced from 2.2 to 1.6) to increase cloudiness. For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.

All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}. In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2. {\color{orange}{Marcus - this is the only part that doesn't talk about CLM/CISM coupling. Move somewhere else and change section name?}}

All grids and dycores simulate the GrIS Surface Mass Balance (SMB). The SMB is the sum of mass accumulation of precipitation and mass loss from ablation. Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation runs off the ice sheet; liquid runoff (which is composed of both rain or ice/snow melt) can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass. These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, using a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB. The area fraction of each EC is derived from a high-resolution GrIS elevation dataset. The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean. Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent. Any snow above the 10-m cap is routed to the ocean as solid runoff.  Refreezing of liquid water within the snowpack is an additional source of ice.  Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation. For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{Muntjewerf_EA_2021_coupling,CISM1,SETAL2019TC,KETAL2020JAMES}. {\color{orange}{Marcus - mention you're not running CISM.}}

@@

We use several observational datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations. We gathered SMB datasets from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain boundaries. The Regional Atmospheric Climate Model (\textit{RACMO}) simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite{NETAL2015TC,NETAL2019SCIENCE}.

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}). The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}. CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite<ESMF;>{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.

@@

>>>>

In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs \cite<see>[and references therein]{W2007JMSJ}.
These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly. While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid. We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.

@@

The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which have large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}. Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c). Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{PL2007JCP,DetAl2012IJHPCA,WETAL2013GMD}. This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement. In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).  However, refinement comes at the cost of a smaller CFL-limited time step in the refined region. (The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.)

@@

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias \cite{P2000GPC,VETAL2018TC}. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low resolution GCMs, and may further degrade the simulated SMB. For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates. The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}. However, unlike GCMs, RCMs are not a freely evolving system and boundary conditions must be prescribed at the lateral boundaries. The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system. In addition, the boundary conditions are derived from a separate host model, which introduces inconsistencies due to differences in model design between the host model and the RCM.

@@

The SE dycore has been an available option since CESM, version 1, but has been under active development ever since. This includes the switch to a dry-mass vertical coordinate \cite{LetAl2018JAMES} and incorporation of an accelerated multi-tracer transport scheme \cite{LTOUNGK2017MWR}, made available in CESM2. This paper documents several additional enhancements to the SE dycore as part of the release of CESM2.2. These include three new VR configurations (Figure~\ref{fig:vr-grids}), two Arctic meshes and a Contiguous United-States mesh ($\tt CONUS$; featured in \citeA{PETAL2020BAMS}). While there are dozens of published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,GetAl2017JAMES,BETAL2019TAC,BETAL2021IJC}, these studies either used development code or collaborated closely with model developers. CESM2.2 is the first code release that contains out of the box VR functionality in CESM.

This study compares the representation of Arctic regions using the SE and FV dycores in CESM2.2 (see description below), as these two dycores treat high latitudes (i.e., the pole problem) in different ways. Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods. Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.

@@

The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores. These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models. This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores. CESM2 runs submitted to the Coupled Model Intercomparison Project Phase 6 \cite{CMIP6} are based on the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.

@@

The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}. The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate. Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong]. A polar filter damps computational instability due to the convergence of meridians, permitting a more practical time step. It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}. The form of the filter is designed to slow down the propagation of large zonal wave-numbers to satisfy the CFL condition of the shortest resolved wave at some reference latitude.

@@

The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups. (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing. A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements. (2) The topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.

@@

We evaluate model simulations on six different grids in this study (Table~\ref{tbl:table1}). The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b). We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMES}.

Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and $\tt CONUS$ (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region. The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution. The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else. The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations using the $\tt CONUS$ grid.

@@

The physics time step depends on the grid resolution. Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $\tt Arctic$ and $\tt Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step used in the $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).

@@

All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}. The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}. Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).

Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore. The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance. We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb\_gamma=0.308\rightarrow0.270$) and also reduced the strength of the damping for horizontal component of turbulent energy ($\tt clubb\_c14=2.2\rightarrow1.6$) to increase cloudiness. For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.

\subsection{Modeled SMB}

All grids and dycores simulate the GrIS Surface Mass Balance (SMB). The SMB is the sum of mass accumulation of precipitation and mass loss from ablation. Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt. Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass. These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.

CLM runs on the same grid as the atmosphere, using a downscaling technique to account for sub-grid variability in SMB. In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB. The area fraction of each EC is derived from a high-resolution GrIS elevation dataset. The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean. Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent. Any snow above the 10-m cap is routed to the ocean as solid runoff.  Refreezing of liquid water within the snowpack is an additional source of ice.  Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation. For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}. {\color{orange}{Marcus - mention you're not running CISM.}}

@@

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}). The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}. CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.

@@

All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}. In CESM terminology, AMIP simulations use the $FHIST$ computational set and run out of the box in CESM2.2.

We use several observational datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations. We gathered SMB datasets from several sources. Regional Atmospheric Climate Model, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain's lateral boundaries. The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite<e.g.,>{LIVVkit,KETAL2020JAMES}.
